Gradual Soundness: Lessons from Static Python

### 2021-12-21 pj-7.1/paper.tex
pj: sec 2.2 draft

\subsubsection{Bytecode Optimization}

TODO

optimizations and bytecodes for: length, subscripting, function dispatch, method dispatch, field load/store

table much later in paper

IN ADDITION to optimizations ...
not sure where this belongs exactly, but we go to a lot of work in the compiler
to intelligently minimize runtime checks by only inserting them when the source
value is of dynamic type (this includes also skipping CHECK\_ARGS if the
function was statically invoked by another static function and thus we know the
compiler checked argument types)

>>>>

\subsubsection{Bytecode Instructions}

Cinder is a bytecode compiler that extends the Python instruction set.
All the standard Python instructions work similarly to Python 3.8;
in other words, an untyped Python program has the same semantics in
Cinder as in the standard interpreter.
The added instructions help either to express runtime checks or to
quickly perform a statically-determined action.

The two main instructions that express run-time checks are \code{CAST}
and \code{CHECK\_ARGS}.
The former checks a value against a type.
The latter is a special cast for functions and methods;
it appears at the top of every function and checks that the
actual parameters match the types of the formal parameters.

A simple optimizing instruction is \code{FAST\_LEN}, which expects
a built-in value\footnote{Because of subtyping, it does not suffice to check
the type. All Python lists support \code{FAST\_LEN} but some subclasses
of the Python list type may not.}
and quickly computes its length.
Another is \code{LOAD\_FIELD}, which scans a fixed offset within an object.
By contrast, the standard Python instruction for loading a field needs to
search a dictionary~(\code{LOAD\_ATTR}).

In addition to speeding code up, the \SP{} compiler takes care
to avoid slowing code down.
It does so by avoiding casts and run-time checks that are certain to succeed.
For example, calls from a typed context to a typed function are statically
checked for compatibility.
There is no need to run \code{CHECK\_ARGS} on typed inputs, so the bytecode
compiler inserts a special call instruction to skip past the check.
